Using Scagnostics to find Interesting Projections of Multivariate Data

### 2018-02-14 Draft/Draft.Rmd
restructuring paper

---
author:
  - name: Ursula Laa
    affiliation: Monash University
    address: >
      First line
      Second line
    email: \email{ursula.laa@monash.edu}
    url: http://rstudio.com
  - name: Second Author
    affiliation: Affiliation
title:
  formatted: "Scagnostics and Projection Pursuit"
  # If you use tex in the formatted title, also supply version without
  #plain:     "A Capitalized Title: Something about a Package foo"
  # For running headers, if needed
  #short:     "\\pkg{foo}: A Capitalized Title"
abstract: >
  The abstract of the article.
keywords:
  # at least one keyword must be supplied
  formatted: [scagnostics, tour, projection pursuit]
  #plain:     [keywords, not capitalized, Java]
preamble: >
  \usepackage{amsmath}
output: rticles::jss_article
---

@@

@@

# Scagnostics measures

@@

# Analysis setup

@@

## Scagnositics of grand tour projections

```{r scagnostics-summary, fig.height= 14, fig.width=9, fig.cap = "(ref:scagnostics-summary)"}
```

@@

```{r scatter-binning, fig.height=2, fig.cap = "(ref:scatter-binning)"}
```

```{r scatter-table}
```

@@

```{r scagnositcs-timing, fig.cap = "(ref:scagnostics-timing)"}
```

@@

We next turn to the question of how we can obtain smooth transitions of the scagnostics measures.
(Below just an example of using ggplot smoothing to display what kind of behaviour we want to find. It follows the general trend without capturing the spikes.)
```{r smoothing, fig.cap = "(ref:smooting)"}
```

@@

```{r guided-tour, include=FALSE}
```
```{r guided-tour-plot, fig.height = 3, fig.cap = "(ref:guided-tour-plot)"}
```

Figure (ref:guided-tour-scatter) compares the scatter plots (left unbinned, right binned) of the start projection, the projection with the lowest index value (t=19 of the interpolated projections) and the final projection (t=28).
```{r guided-tour-scatter, fig.height = 3, fig.cap = "(ref:guided-tour-scatter)"}
```

@@

```{r mbgraphics-indices, fig.height=3, fig.cap="(ref:mbgraphics-indices)"}
```
```{r guided-spline2d, include=FALSE}
```
```{r guided-spline2d-plot, fig.height=3, fig.cap="(ref:guided-spline2d-plot)"}
```
```{r guided-dcor2d, include=FALSE}
```
```{r guided-dcor2d-plot, fig.height=3, fig.cap="(ref:guided-dcor2d-plot)"}
```

@@

```{r gminus2-box, fig.height=3}
```

```{r guided-dcor2d-no-gminus2, include=FALSE}
```
```{r guided-dcor2d-plot-no-gminus2, fig.height=3, fig.cap="(ref:guided-dcor2d-plot)"}
```

>>>>

---
author:
  - name: Ursula Laa
    affiliation: Monash University
    address: >
      First line
      Second line
    email: \email{ursula.laa@monash.edu}
    url: http://rstudio.com
  - name: Second Author
    affiliation: Affiliation
title:
  formatted: "Using Scagnostics to find Interesting Projections of Multivariate Data"
  # If you use tex in the formatted title, also supply version without
  #plain:     "A Capitalized Title: Something about a Package foo"
  # For running headers, if needed
  #short:     "\\pkg{foo}: A Capitalized Title"
abstract: >
  The abstract of the article.
keywords:
  # at least one keyword must be supplied
  formatted: [scagnostics, tour, projection pursuit, multivariate data, statistical graphics, data visualisation]
  #plain:     [keywords, not capitalized, Java]
preamble: >
  \usepackage{amsmath}
output: rticles::jss_article
---

@@

*** Motivating example

```{r motivation}
library(mvtnorm)
x1 <- rmvnorm(100, mean=c(0,0), 
             sigma=matrix(c(1, 0.8, 0.8, 1), ncol=2, byrow=T))
x2 <- rmvnorm(100, mean=c(3,0), 
             sigma=matrix(c(1, 0.8, 0.8, 1), ncol=2, byrow=T))
x3 <- rmvnorm(100, mean=c(0,3), 
             sigma=matrix(c(1, 0.8, 0.8, 1), ncol=2, byrow=T))
df <- data.frame(x1=rnorm(300),
                 x2=rnorm(300),
                 x3=rnorm(300),
                 x4=rnorm(300), 
                 x5=scale(c(x1[,1], x2[,1], x3[,1])), 
                 x6=scale(c(x1[,2], x2[,2], x3[,2])), 
                 cl=c(rep("A",100), rep("B", 100), rep("C",100)))
ggplot(df, aes(x=x5, y=x6, colour=cl)) + geom_point() + theme(aspect.ratio=1)
```

```{r}
library(tourr)
quartz()
col <- c(rep("#f0a800", 100), rep("#486030", 100), rep("#c03018", 100))
animate(df[,1:6], guided_tour(lda_pp(df$cl)),
        display_xy(axes="bottomleft", col=col))
```

```{r}
library(geozoo)

```

*** Explanation of tour algorithm and projection pursuit and software

*** Explanation of scagnostics and software

@@

# Investigation of scagnostics

## Scagnostic discreteness

*** temporal trace of scagnostics showing discreteness, with projections illustrating little difference

*** katrin's work

@@

## Speed

# Modification to scagnostics

## Smoothing

## Jittering

## Tour optimisation

# Scagnostics indices

## Individual

## Two group differences

## Multiple group differences

@@

## Scagnostics of grand tour projections

```{r scagnostics-summary, fig.height= 14, fig.width=9, fig.cap = "(ref:scagnostics-summary)", eval=FALSE}
```

@@

```{r scatter-binning, eval=FALSE, fig.height=2, fig.cap = "(ref:scatter-binning)"}
```

```{r scatter-table, eval=FALSE}
```

@@

```{r scagnositcs-timing, eval=FALSE, fig.cap = "(ref:scagnostics-timing)"}
```

@@

We next turn to the question of how we can obtain smooth transitions of the scagnostics measures.
(Below just an example of using ggplot smoothing to display what kind of behaviour we want to find. It follows the general trend without capturing the spikes.)
```{r smoothing, eval=FALSE, fig.cap = "(ref:smooting)"}
```

@@

```{r guided-tour, eval=FALSE, include=FALSE}
```
```{r guided-tour-plot, eval=FALSE, fig.height = 3, fig.cap = "(ref:guided-tour-plot)"}
```

Figure (ref:guided-tour-scatter) compares the scatter plots (left unbinned, right binned) of the start projection, the projection with the lowest index value (t=19 of the interpolated projections) and the final projection (t=28).
```{r guided-tour-scatter, eval=FALSE, fig.height = 3, fig.cap = "(ref:guided-tour-scatter)"}
```

@@

```{r mbgraphics-indices, eval=FALSE, fig.height=3, fig.cap="(ref:mbgraphics-indices)"}
```
```{r guided-spline2d, eval=FALSE, include=FALSE}
```
```{r guided-spline2d-plot, eval=FALSE, fig.height=3, fig.cap="(ref:guided-spline2d-plot)"}
```
```{r guided-dcor2d, eval=FALSE, include=FALSE}
```
```{r guided-dcor2d-plot, eval=FALSE, fig.height=3, fig.cap="(ref:guided-dcor2d-plot)"}
```

@@

```{r gminus2-box, eval=FALSE, fig.height=3}
```

```{r guided-dcor2d-no-gminus2, eval=FALSE, include=FALSE}
```
```{r guided-dcor2d-plot-no-gminus2, eval=FALSE, fig.height=3, fig.cap="(ref:guided-dcor2d-plot)"}
```

### 2018-02-14 Draft/Draft.tex
restructuring paper

\author{
Ursula Laa\\Monash University \And Second Author\\Affiliation
}
\title{Scagnostics and Projection Pursuit}

@@

\Keywords{scagnostics, tour, projection pursuit}

@@

\subsection{Scagnositics of grand tour
projections}\label{scagnositics-of-grand-tour-projections}

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=\textwidth]{figure/scagnostics-summary-1}

}

\caption[(ref:scagnostics-summary)]{(ref:scagnostics-summary)}\label{fig:scagnostics-summary}
\end{figure}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=\textwidth]{figure/scatter-binning-1} \includegraphics[width=\textwidth]{figure/scatter-binning-2} \includegraphics[width=\textwidth]{figure/scatter-binning-3}

}

\caption[(ref:scatter-binning)]{(ref:scatter-binning)}\label{fig:scatter-binning}
\end{figure}
\end{CodeChunk}

\begin{CodeChunk}

\begin{CodeOutput}
                   s1          s2          s3
Outlying  0.275676305 0.163548110 0.270907044
Skewed    0.549038842 0.575422054 0.583569727
Clumpy    0.004123711 0.008163265 0.004175365
Sparse    0.038765964 0.042779580 0.040551326
Striated  0.059405941 0.060000000 0.037433155
Convex    0.611714549 0.602591732 0.546399191
Skinny    0.483884014 0.495557112 0.564627720
Stringy   0.424567811 0.368764181 0.329434667
Monotonic 0.077330543 0.094217913 0.103036354
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=\textwidth]{figure/scagnositcs-timing-1}

}

\caption[(ref:scagnostics-timing)]{(ref:scagnostics-timing)}\label{fig:scagnositcs-timing}
\end{figure}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=\textwidth]{figure/smoothing-1}

}

\caption[(ref:smooting)]{(ref:smooting)}\label{fig:smoothing}
\end{figure}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=\textwidth]{figure/guided-tour-plot-1} \includegraphics[width=\textwidth]{figure/guided-tour-plot-2}

}

\caption[(ref:guided-tour-plot)]{(ref:guided-tour-plot)}\label{fig:guided-tour-plot}
\end{figure}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=\textwidth]{figure/guided-tour-scatter-1} \includegraphics[width=\textwidth]{figure/guided-tour-scatter-2} \includegraphics[width=\textwidth]{figure/guided-tour-scatter-3}

}

\caption[(ref:guided-tour-scatter)]{(ref:guided-tour-scatter)}\label{fig:guided-tour-scatter}
\end{figure}
\end{CodeChunk}

@@

Additions to the standard scagnostics measures have been implemented in
the package mbgraphics, based on splines (spline2d) and distance
correlation (dcor2d). Give definitions here.. We repeat the study from
before, we first test in Figure ADDREF the behaviour of the index values
over 100 interpolated projections that are randomly selected via the
grand tour algorithm (the same projections as considered above). Note
that here the measures are calculated on non-binned data. The general
behaviour seems similar to the scagnostics monotonic index. The observed
behaviour is smooth (much in contrast to the behaviour of the
scagnostics measures) and we next test the behaviour in a guided tour,
comparing again also the first and final projection. (which index to
select, will try both below\ldots{}) It seems they both maximise
correlation, this might not be too useful?

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=\textwidth]{figure/mbgraphics-indices-1}

}

\caption[(ref:mbgraphics-indices)]{(ref:mbgraphics-indices)}\label{fig:mbgraphics-indices}
\end{figure}
\end{CodeChunk}

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics[width=\textwidth]{figure/guided-spline2d-plot-1} \includegraphics[width=\textwidth]{figure/guided-spline2d-plot-2} \includegraphics[width=\textwidth]{figure/guided-spline2d-plot-3}

}

\caption[(ref:guided-spline2d-plot)]{(ref:guided-spline2d-plot)}\label{fig:guided-spline2d-plot}
\end{figure}
\end{CodeChunk}

\begin{CodeChunk}

\begin{CodeOutput}
            [,1]        [,2]
[1,]  0.47851413 -0.54459454
[2,]  0.01632788 -0.00642453
[3,]  0.20956852 -0.15615632
[4,] -0.21613392  0.22043552
[5,] -0.26952654  0.31203441
[6,] -0.27310129  0.26333539
[7,] -0.72999721 -0.68094632
\end{CodeOutput}
\begin{figure}

{\centering \includegraphics[width=\textwidth]{figure/guided-dcor2d-plot-1} \includegraphics[width=\textwidth]{figure/guided-dcor2d-plot-2} \includegraphics[width=\textwidth]{figure/guided-dcor2d-plot-3}

}

\caption[(ref:guided-dcor2d-plot)]{(ref:guided-dcor2d-plot)}\label{fig:guided-dcor2d-plot}
\end{figure}
\end{CodeChunk}

>>>>

\author{
Ursula Laa\\Monash University \And Second Author\\Affiliation
}
\title{Using Scagnostics to find Interesting Projections of Multivariate Data}

@@

\Keywords{scagnostics, tour, projection pursuit, multivariate data, statistical graphics, data visualisation}

@@

\subsection{Scagnostics of grand tour
projections}\label{scagnostics-of-grand-tour-projections}

@@

@@

@@

@@

@@

@@

Additions to the standard scagnostics measures have been implemented in
the package mbgraphics, based on splines (spline2d) and distance
correlation (dcor2d). Give definitions here.. We repeat the study from
before, we first test in Figure ADDREF the behaviour of the index values
over 100 interpolated projections that are randomly selected via the
grand tour algorithm (the same projections as considered above). Note
that here the measures are calculated on non-binned data. The general
behaviour seems similar to the scagnostics monotonic index. The observed
behaviour is smooth (much in contrast to the behaviour of the
scagnostics measures) and we next test the behaviour in a guided tour,
comparing again also the first and final projection. (which index to
select, will try both below\ldots{}) It seems they both maximise
correlation, this might not be too useful? Seems like {[},1{]} basically
-{[},2{]} in the final projection (except for {[}7,{]})? probably we see
only structure of gminus2 variable, which has very narrow peak, only few
outlying points is there a better way of transforming it to spread out
the structure?
